---
title: "Assignment: Chapter 1 HW"
author: "Type Your Name Here"
date: "Type The Date Here"
output: html_document
---

### Problem 5

In a population of size $N$, the probability of any subset of size $n$ being chosen is $1/{N \choose n}$. 

a) Show that this implies that any one person in the population has a $n/N$ probability of being chosen in a sample. 

b) Then, in particular, show that every person in the population has the same probability of being chosen.

```{r}
#R CODE HERE
```

### Problem 6

A typical Gallup poll surveys about $n$ = 1000 adults. Suppose the sampling frame contains 100 million adults (including you). Now, select a random sample of 1000 adults.

a) What is the probability that you will be in the sample?

b) Now suppose that 2000 such samples are selected, each independent of the others. What is the probability that you will *not* be in any of the samples?

c) How many samples must be selected for you to have a 0.5 probability of being in at least one sample?

```{r}
#R CODE HERE
```

### Appendix

#### Inline Equation Example

${10 \choose 2}=\frac{10!}{8!2!}=\frac{10 \times 9 \times 8 \times 7 \times 6 \times 5 \times 4\times 3 \times 2 \times 1}{(8 \times 7 \times 6 \times 5 \times 4\times 3 \times 2 \times 1)(2 \times 1)}=\frac{10 \times 9}{2 \times 1}=\frac{90}{2}=45$

#### Displayed Equation Example

$${10 \choose 2}=\frac{10!}{8!2!}=\frac{10 \times 9 \times 8 \times 7 \times 6 \times 5 \times 4\times 3 \times 2 \times 1}{(8 \times 7 \times 6 \times 5 \times 4\times 3 \times 2 \times 1)(2 \times 1)}=\frac{10 \times 9}{2 \times 1}=\frac{90}{2}=45$$