---
title: "Assignment: Chapter 1 HW"
author: "Callum Feher"
date: "1/26/15"
output: html_document
---

### Problem 5

In a population of size $N$, the probability of any subset of size $n$ being chosen is $1/{N \choose n}$. 

a) Show that this implies that any one person in the population has a $n/N$ probability of being chosen in a sample. 

$P(n)$ of any one person in the population will be picked for the sample meaning that $1/{N \choose n}$ = $n/N$. So starting with $1/{N \choose n}$ this can be expanded out to $1/{\frac{(N-n)!N}{n!(N-n)}}$ since n is any one person in the population $n! = n$ therefore $1/{\frac{(N-n)!N}{n!(N-n)}}$ can be rewritten as $1/{\frac{N}{n}}$ which ultimately is $n/N$.

b) Then, in particular, show that every person in the population has the same probability of being chosen.

Since $1/{N \choose n}$ is essentially $n/N$ for an individual being chosen in a population it is also the same for a sample size $n$ in the same population.

### Problem 6

A typical Gallup poll surveys about $n$ = 1000 adults. Suppose the sampling frame contains 100 million adults (including you). Now, select a random sample of 1000 adults.

a) What is the probability that you will be in the sample?

```{r}
p <- 1000/(100*10^6)
p
```

b) Now suppose that 2000 such samples are selected, each independent of the others. What is the probability that you will *not* be in any of the samples?

```{r}
(1-p)^2000
```

### Appendix

#### Inline Equation Example

${10 \choose 2}=\frac{10!}{8!2!}=\frac{10 \times 9 \times 8 \times 7 \times 6 \times 5 \times 4\times 3 \times 2 \times 1}{(8 \times 7 \times 6 \times 5 \times 4\times 3 \times 2 \times 1)(2 \times 1)}=\frac{10 \times 9}{2 \times 1}=\frac{90}{2}=45$

#### Displayed Equation Example

$${10 \choose 2}=\frac{10!}{8!2!}=\frac{10 \times 9 \times 8 \times 7 \times 6 \times 5 \times 4\times 3 \times 2 \times 1}{(8 \times 7 \times 6 \times 5 \times 4\times 3 \times 2 \times 1)(2 \times 1)}=\frac{10 \times 9}{2 \times 1}=\frac{90}{2}=45$$