---
title: "Homogenindependence"
author: "Callum Feher"
date: "4/15/15"
output: html_document
---

### Problem 1
Ho = Is that the personality types between gender are the same

Ha = Is that the personality types between gender are different

First define the chi function.
```{r}
chisq<-function(Obs)
{ 
  Exp <- outer(rowSums(Obs),colSums(Obs))/sum(Obs)
  sum((Obs-Exp)^2/Exp)
}

```

Load the data set.
```{r}
STT3850 <- read.csv(file=url("http://www1.appstate.edu/~thomleyje/R-datafiles/PersonalitySTT3850.csv"))
```

Select the two variables we want to test and make a table.Then use the function to calculate the chi-square test statistic (obs.diff)

```{r}
Observed <- table(STT3850$Gender,STT3850$Type)
Observed
obs.diff <- chisq(Observed)
```

Permutation sampling simulation, start by selecting the two variables that are being analysed then run the simulation.

```{r}
set.seed(0)
Sex    <- subset(STT3850, select=Gender, drop=T)
Type <- subset(STT3850, select=Type,drop=T)

N <- 10^4-1
random.diff<-numeric(N)
for (i in 1:N)
{
  TYPE.perm <-sample(Type)
  RANDOM.table <- table(Sex, TYPE.perm)
  random.diff[i]<-chisq(RANDOM.table)
}
```

Graphing the results of the simulation and viewing the expected values table.
```{r}
Expected <- outer(rowSums(Observed),colSums(Observed))/sum(Observed)
Expected

hist(random.diff, prob=T,
     main="Null Distribution of Chi-Square Test Statistic",
     xlab="randomly generated chi-square statistics", 
     xlim=c(0,max(max(random.diff),obs.diff)),
     ylim=c(0,1))
abline(v=obs.diff,
       col="blue",
       lty=5)  
```

last but not least the p value 

```{r}
print(paste("The chi-square test statistic is ", round(obs.diff,4), ".", sep=""),quote=F)
print(paste("The p-value for the test of independence is ", format((sum(random.diff >= obs.diff) + 1)/(N + 1),scientific=F), ".", sep=""),quote=F)
```

Rejecting the Ho is the as the p-value of the data is less than significance value of 0.10. Meaning the personality type is different by gender for the data set STT3850.

Notes:
I notices that the expected value table and the observed table had very similar proportions to each other.

Some personality types had a majority of one gender in them.

### Problem 2

The other variable in the data set that could have an influence on the differences in personality type by gender is possible the major each of the students are in.


### Problem 3
The error message is due to the small values in some of the cells in the Observed table shown in problem 1.
```{r}
chisq.test(table(STT3850$Gender,STT3850$Type),correct=F)

```

